Remote sensing the gas transfer velocity using SWOT observations and no in situ information

### 2020-12-09 Outline.Rmd
new model runs

---
title: "Remote sensing the gas transfer velocity using SWOT observations"
author: "Craig Brinkerhoff"
date: "`r Sys.Date()`"
output:
 # pdf_document:
#    latex_engine: xelatex
  html_document: default
---

@@

1)	CO2 emissions from rivers are significant (duh)
2)	Estimating this flux is done by estimating in-stream CO2 and the gas transfer velocity, or ‘evasion rate’ or k600 of the river channel
3)	K600 is incredibly cumbersome to directly measure, and impractical when performing upscaling to entire river networks
4)	Previous work has stressed that this is largely a f(turbulence) driven by river hydraulics/geomorphology
    a.	Empirical scaling functions have been developed to use hydraulics to model k600 for entire river systems but are prone to significant uncertainties (Raymond et al. 2012)
    b.	They are also fundamentally dependent on modeled or measured flow velocities which themselves are prone to significant uncertainties (or poorly informed hydraulic geometry models…)
    c.	What is needed is a simple way to directly observe the gas transfer velocity at scale and high temporal resolution without needing explicit flow velocity estimates. We argue that the upcoming NASA SWOT mission provides exactly the data required to make this a reality and remotely sense (RS) k600 at unprecedented temporal resolution

@@

## Goal 2
1) We propose the BIGER (Bayesian Inference/Inversion of Gas Evasion Rate) algorithm to remotely sense k600 from river width/height alone.
2) We implement the k600~eD model from Q1 as a Bayesian likelihood function following methods developed for ungauged RSQ (Hagemann et al. 2017). Below is the derivation of the likelihood function used in the Bayesian inference model.
	  a. $k_{600}=a(gS[\frac{1}{n} (A_{0}+dA)^{2/3} W^{-2/3} S^{1/2}])^{b}$
	  b. $k_{600}(\frac{1}{n}(A_0+dA)^{2/3})^{-b}=a(g W^{2/3}S^{3/2})^{b}$
	    i. Model parameters are now on left-hand side (i.e. need priors) and the data (i.e. SWOT observables) are on right-hand side. This allows us to, given the constant rhs value, sample from the left in the same manner that a traditional Bayesian sampling model samples from a set of model parameters (see Hagemann etal 2017 & Durand etal 2014). Therefore, the 'model parameters' are k600, n, and A0. In a Bayesian framework, these need priors assigned to them.
	    ii. Brinkerhoff etal 2020 priors are used for n and A0. k600 priors are obtained similarly using Ulseth etal 2019 as the training data. <b> ELABORATE </b>
3) For the sake of validating the algorithm, we assume that the scaling model is ‘truth’ and see if we can infer these values using only SWOT observables (river width and height/slope).
    a. Therefore, in the validation setup, posterior uncertainty is only Manning’s uncertainty (0.25) and not the total model uncertainty obtained in the MC simulations earlier. However, this total uncertainity would be implemented if running the algorithm for general science objectives.
    b. We validated on 10-day pepsi 2 data. Plotted are means and 95% CIs of the posterior k600 values versus the ‘observed’ k600 calculated from observed velocity and slope for all timesteps across all 32 rivers

![**Figure 4** *Validation of remote sensing algorithm for 32 rivers with 10day sampling intervals*](outputs//validation//validation.jpg)

4. Below are boxplots of performance metrics for the 32 rivers, i.e. how well do individual rivers do? NRMSE, RRMSE, and rBIAS are optimally zero while KGE is optimally 1. A KGE greater than -0.41 indicates our model outperforms a mean estimate, akin to an NSE of 0 for the hydrologists out there (Knoben etal 2019). Also see Knoben et al (2019) for why NSE completely blows up for k600.
    a. The result is great KGEs, NRMSEs, and RRMSEs (espically for aboslutely no in situ information) however a pretty high rBIAS. This bias is borne out in the scatterplot earlier.

@@

>>>>

---
title: "Remote sensing the gas transfer velocity using SWOT observations and no in situ information"
author: "Craig Brinkerhoff"
date: "`r Sys.Date()`"
output:
 # pdf_document:
#    latex_engine: xelatex
  html_document: default
---

@@

1)	CO2 emissions from rivers are significant (duh)
2)	Estimating this flux is done by estimating in-stream CO2 and the gas transfer velocity, or ‘evasion rate’ or k600 of the river channel
3)	K600 is incredibly cumbersome to directly measure, and impractical when performing upscaling to entire river networks
4)	Previous work has stressed that this is largely a f(turbulence) driven by river hydraulics/geomorphology
    a.	Empirical scaling functions have been developed to use hydraulics to model k600 for entire river systems but are prone to significant uncertainties (Raymond et al. 2012)
5) Further, in situ data for hydraulics is often unavaible (esp. in remote rivers) and so k models are at the mercy of modeled hydraulics, themselves highly uncertain.
6)	What is needed is a simple way to observe the gas transfer velocity at large scales and high temporal resolution without relying on situ flow velocity data or modeled velocities. We argue that the upcoming SWOT mission provides exactly the data required to make this a reality (at least in large rivers) and remotely sense (RS) k600 as a function of river hydraulics.

@@

## Goal 2
1) We propose the BIGER (Bayesian Inference/Inversion of the Gas Evasion Rate) algorithm to remotely sense k600 from river width/height alone.
2) We implement the k600~eD model from Q1 as a Bayesian likelihood function following methods developed for ungauged RSQ (Hagemann et al. 2017). Below is the derivation of the likelihood function used in the Bayesian inference model.
	  a. $k_{600}=a(gS[\frac{1}{n} (A_{0}+dA)^{2/3} W^{-2/3} S^{1/2}])^{b}$
	  b. $k_{600}(\frac{1}{n}(A_0+dA)^{2/3})^{-b}=a(g W^{2/3}S^{3/2})^{b}$
	    i. Model parameters are now on left-hand side (i.e. need priors) and the data (i.e. SWOT observables) are on right-hand side. This allows us to, given the constant rhs value, sample from the left in the same manner that a traditional Bayesian sampling model samples from a set of model parameters (see Hagemann etal 2017 & Durand etal 2014). Therefore, the 'model parameters' are k600, n, and A0. In a Bayesian framework, these need priors assigned to them.
	    ii. Brinkerhoff etal 2020 priors are used for n and A0. k600 priors are obtained similarly using Ulseth etal 2019 as the training data. <b> ELABORATE </b>
3) For the sake of validating the algorithm, we assume that the scaling model is ‘truth’ and see if we can infer these values using only SWOT observables (river width and height/slope).
    a. Therefore, in the validation setup, posterior uncertainty is only Manning’s uncertainty (0.25) and not the total model uncertainty obtained in the MC simulations earlier. However, this total uncertainity would be implemented if running the algorithm for general science objectives.
    b. We validated on 10-day pepsi 2 data. Plotted are means and 95% CIs of the posterior k600 values versus the ‘observed’ k600 calculated from observed velocity and slope for all timesteps across all 32 rivers

![**Figure 4** *Validation of remote sensing algorithm for 32 rivers with 10day sampling intervals. Black bars are 95% CIs for the modeled values. Black line is linear regression and dashed grey line is 1:1 line.*](outputs//validation//validation.jpg)

4. Below are boxplots of performance metrics for the 32 rivers, i.e. how well do individual rivers do? NRMSE, RRMSE, and rBIAS are optimally zero while KGE is optimally 1. A KGE greater than -0.41 indicates our model outperforms a mean estimate, akin to an NSE of 0 (Knoben etal 2019). Also see Knoben et al (2019) for why NSE completely blows up for k600.

@@

##References
Durand, M., Neal, J., Rodríguez, E., Andreadis, K. M., Smith, L. C., & Yoon, Y. (2014). Estimating reach-averaged discharge for the River Severn from measurements of river water surface elevation and slope. Journal of Hydrology, 511, 92–104. https://doi.org/10.1016/j.jhydrol.2013.12.050

Engram, M., Walter Anthony, K. M., Sachs, T., Kohnert, K., Serafimovich, A., Grosse, G., & Meyer, F. J. (2020). Remote sensing northern lake methane ebullition. Nature Climate Change, 1–7. https://doi.org/10.1038/s41558-020-0762-8

Hagemann, M. W., Gleason, C. J., & Durand, M. T. (2017). BAM: Bayesian AMHG-Manning Inference of Discharge Using Remotely Sensed Stream Width, Slope, and Height: BAM FLOW USING STREAM WIDTH SLOPE HEIGHT. Water Resources Research, 53(11), 9692–9707. https://doi.org/10.1002/2017WR021626

Knoben, W. J. M., Freer, J. E., & Woods, R. A. (2019). Technical note: Inherent benchmark or not? Comparing Nash–Sutcliffe and Kling–Gupta efficiency scores. Hydrology and Earth System Sciences, 23(10), 4323–4331. https://doi.org/10.5194/hess-23-4323-2019
